GLMM

### 2024-06-20 Code/GLMM.Rmd
Added womens and dash analyses

times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
with2022 <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
without2022 <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Venue != 2022)
```

@@

### 20
This code creates the venue effect only models.
```{r}
dash_no_20222022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_no_2022, 
                                       family = Gamma(link = "log"))
dash_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_2022, 
                                   family = Gamma(link = "log"))
```

@@

### 23
Now we test
```{r}
sim_two_effect_time(dash_no_2022_two_effect_gamma_log,
                    dash_no_2022_one_gamma_log, 0.1)
sim_two_effect_time(dash_2022_two_effect_gamma_log,
                    dash_2022_one_gamma_log, 0.1)
```

>>>>

times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
with2022 <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ' & Event == "Hurdles")
without2022 <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Venue != 2022 & Event == "Hurdles")
```

@@

### 20
This code creates the venue effect only models.
```{r}
dash_no_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_no_2022, 
                                       family = Gamma(link = "log"))
dash_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_2022, 
                                   family = Gamma(link = "log"))
```

@@

### 23
Now we test
```{r}
sim_two_effect_time(dash_no_2022_two_effect_gamma_log,
                    dash_no_2022_one_gamma_log, 0.1)
sim_two_effect_time(dash_2022_two_effect_gamma_log,
                    dash_2022_one_gamma_log, 0.1)
```
## 24
In response to reviewer 2's comments that questioned the probability of observing
the venue effect for 2022 we perform the following analysis. We store the venue
effect for 2022 in `effect2022` and we store the standard deviation in `sd2022`
```{r}
library(tidyverse)
ranef_results <- data.frame(ranef(with2022_two_effect_gamma_log))
venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
venues$year <- as.character(venues$grp)
venues$year <- as.numeric(venues$year)
effect2022 <- venues |> filter(grp == 2022) |> select(condval) |> pull()
sd2022 <- as.data.frame(VarCorr(with2022_two_effect_gamma_log)) |> filter(grp == "Venue") |> select(sdcor)|> pull()

```
Now we can calculate how extreme this value is
```{r}
pnorm(effect2022, mean = 0, sd = sd2022)
```

## Now we look at the 200 meter dash results
```{r}
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

```{r}
library(tidyverse)
dash200_2022 <- times |>
  filter(Event == "200 Dash") |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

dash200_no_2022 <- times |>
  filter(Event == "200 Dash") |>
  filter(Venue != 2022) |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

```{r}
library(lme4)
dash200_no_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash200_no_2022, 
                                    family = Gamma(link = "log"))
dash200_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash200_2022, 
                                family = Gamma(link = "log"))

dash200_no_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                                   data = dash200_no_2022,
                                   family = Gamma(link = "log"))
dash200_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = dash200_2022,
                               family = Gamma(link = "log"))
```
Now we test
```{r}
sim_two_effect_time(dash200_no_2022_two_effect_gamma_log,
                    dash_no_2022_one_gamma_log, 0.1)
sim_two_effect_time(dash200_2022_two_effect_gamma_log,
                    dash_2022_one_gamma_log, 0.1)
```
These are low probabilities indicating that a reaction time below 0.1 seconds
is unlikely for the 200 meter dash.
```{r}
library(ggplot2)
ggplot(dash200_2022, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.25),
                            breaks = seq(0.1, 0.25, 0.05)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
```

## We will compare our results to the womens data starting with hurdles and then dash

```{r}
library(tidyverse)
women_hurdles <- times |>
  filter(Event == "100 Hurdles") |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "F") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

women_dash <- times |>
  filter(Event == "100 Dash") |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "F") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

```{r}
library(ggplot2)
ggplot(women_hurdles, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.25),
                            breaks = seq(0.1, 0.25, 0.05)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(2000,2024),
                                              breaks = seq(2001, 2024, 2))
```
```{r}
library(ggplot2)
ggplot(women_dash, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.25),
                            breaks = seq(0.1, 0.25, 0.05)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
```
